The Morality of Arbitrage - Some Half-Baked Ideas

### 2016-09-23 Slides/morality-slides.Rmd
Fixed a few errors.

---
title: "The Morality of Arbitrage - Some Half-Baked Ideas"
author: "Tyler J. Brough"
date: "September 16, 2016"
output: ioslides_presentation
---

@@

@@

> _"I would put the point more bluntly. The market for ideas is the market in which the intellectual conducts his trade. The explanation of the paradox is self-interest and self-esteem. Self-esteem leads the intellectuals to magnify the importance of their own market. That others should be regulated seems naturally, particularly as many of the intellectuals see themselves as doing the regulating. But self-interest combines with self-esteem to ensure that, while others are regulated, regulation should not apply to them."_

@@

> _"I do not believe that we will be able to form a judgement in which we can have any confidence unless we abandon the present ambivalence about the performance of government in the two markets and adopt a more consistent view"_

@@

>- My finance colleagues will understand de Finetti's lottery tickets as contingent claims.
>- In fact, de Finetti already invented much of what we thing of as modern finance. [But no one really knew about it!](https://www.openstarts.units.it/dspace/bitstream/10077/3969/1/Pressacco%20Ziani%20Atti%20Economia%20e%20Incertezza.pdf)

@@

According to *The Cambridge Dictionary of Philosophy* at Dutch book argument is the following:

@@

## Here's An Easier Version

From the great comic strip [xkcd](https://xkcd.com/1132/)

```{r, echo=FALSE, out.width = "600px", out.height="450px"}
knitr::include_graphics("images/xkcd.png")
```

@@

> _"In this paper it will be argued that decision theory, game theory, and market theory can be placed on a more secure and parsimonious footing by appeal to the principle of_ ___no arbitrage___ _rather than the usual ordering axioms of preferences. The arbitrage princile already has a long history: in the literature on Bayesian statistics and decision theory, it was introduced by de Finetti (1937) as a basis for defining subjective probability... In the finance literature, the arbitrage principle explicitly underlies the capital structure theorem (Modigliani & Miller 1958), the valuation of cash flows (Ross 1978), and the pricing of assets and options (Black and Scholes 1973, Merton 1973, Ross 1976). No-arbitrage is central to classical welfare economics, where it goes by the name of 'Pareto optimality.'_

@@

>- So the Dutch book, or equivalently the no-arbitrage principle becomes a unifying theme. 
>- We can now explicitly make Coase's argument as a Dutch book argument and point out the incoherence

@@

>- Hayek points out that market processes are knowledge processes. 
>- It should seem very natural to economists to treat the market for ideas and the market for goods similarly. The are about the same thing!
>- Thus, the process of scholarship and science are very much like market process for goods.

Additionally,

@@

## Thomas Sowell Quotation

@@

>- The process of scholarship is generally accepted as an inherently moral pursuit!
>- Scholarship and science are about important ideas and beliefs
>- Economics and finance are about more mundane information processes, but certainly no less vital. Probably much more important to our daily lives!

@@

1. The pursuit of knowledge (_intellectual arbitrage_) is inherently moral

@@

2. The pursuit of capitalist enterprise (_entrepreneurial arbitrage_) is inherently immoral

without leaving yourself open to a massive Dutch book, for they are just two different examples of the very same thing!

@@

>- I am a bit embarrased to admit that until very recently I was not aware of Israel Kirzner's very important work on entrepreneurial arbitrage.
>- Dutch books in the election cycle? Like Thomas Sowell once said, "I feel like a mosquito at a nudist collony. Where do I begin?!"
>- Ben thinks that one of the problems is the lack of arbiragable positions. 
>- We all have the same vote, so there is no one person who can really move the market with superior knowledge. What we get is mass confusion!

>>>>

---
title: "The Morality of Arbitrage - Some Half-Baked Ideas"
author: "Tyler J. Brough"
date: "September 23, 2016"
output: ioslides_presentation
---

@@

The freedom of speech, a free press, religious liberty are so ingrained in our culture. Their morality is based on a generally held and accepted ethical foundation.

@@

> _"I would put the point more bluntly. The market for ideas is the market in which the intellectual conducts his trade. The explanation of the paradox is self-interest and self-esteem. Self-esteem leads the intellectuals to magnify the importance of their own market. That others should be regulated seems natural, particularly as many of the intellectuals see themselves as doing the regulating. But self-interest combines with self-esteem to ensure that, while others are regulated, regulation should not apply to them."_

@@

> _"I do not believe that we will be able to form a judgement in which we can have any confidence unless we abandon the present ambivalence about the performance of government in the two markets and adopt a more consistent view."_

@@

>- My finance colleagues will understand de Finetti's lottery tickets as contingent claims.
>- In fact, de Finetti already invented much of what we think of as modern finance. [But no one really knew about it!](https://www.openstarts.units.it/dspace/bitstream/10077/3969/1/Pressacco%20Ziani%20Atti%20Economia%20e%20Incertezza.pdf)

@@

According to *The Cambridge Dictionary of Philosophy* a Dutch book argument is the following:

@@

## xkcd On Dutch Books

```{r, echo=FALSE, out.width = "600px", out.height="490px"}
knitr::include_graphics("images/xkcd.png")
```

Here is the [link.](https://xkcd.com/1132/)

@@

> _"In this paper it will be argued that decision theory, game theory, and market theory can be placed on a more secure and parsimonious footing by appeal to the principle of_ ___no arbitrage___ _rather than the usual ordering axioms of preferences. The arbitrage princile already has a long history: in the literature on Bayesian statistics and decision theory, it was introduced by de Finetti (1937) as a basis for defining subjective probability... In the finance literature, the arbitrage principle explicitly underlies the capital structure theorem (Modigliani & Miller 1958), the valuation of cash flows (Ross 1978), and the pricing of assets and options (Black and Scholes 1973, Merton 1973, Ross 1976). No-arbitrage is central to classical welfare economics, where it goes by the name of 'Pareto optimality.'"_

@@

>- So the Dutch book, or equivalently the no-arbitrage principle, becomes a unifying theme. 
>- We can now explicitly make Coase's argument as a Dutch book argument and point out the incoherence.

@@

>- Hayek points out that market processes are knowledge processes. 
>- It should seem very natural to economists to treat the market for ideas and the market for goods similarly. They are two different markets, but the same market process!
>- Thus, the process of scholarship and science are very much like the market process for goods.

Additionally

@@

## Thomas Sowell Quotations

@@

>- The process of scholarship is generally accepted as an inherently moral pursuit!
>- Scholarship, science, and journalism are about important ideas and beliefs.
>- Economics and finance are about more mundane information processes, but certainly no less vital. Probably much more important to our daily lives!

@@

  $(1)$ The pursuit of knowledge (_intellectual arbitrage_) is inherently moral

@@

  $(2)$ The pursuit of capitalist enterprise (_entrepreneurial arbitrage_) is inherently immoral

without leaving oneself open to a massive Dutch book. They are just two different examples of the very same thing!

@@

>- I am a bit embarrased to admit that until very recently I was not aware of Israel Kirzner's very important work on entrepreneurial arbitrage.
>- Dutch books in the election cycle? Like Thomas Sowell once said, "I feel like a mosquito at a nudist collony. Where do I begin?!"
>- Ben thinks that one of the problems is the lack of arbitragable positions. 
>- We all have the same vote, so there is no one person who can really move the market with superior knowledge. What we get is mass confusion! 
>- Once you start seeing the world this way you will see Dutch books everywhere. Especially on Facebook.
